---
title: "Stock Market Prediction"
author: "Aymen Rumi"
date: "5/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(ISLR)
library(MASS)
library(class)

knitr::opts_chunk$set(comment = NA)

```

## Overview

We will try to build a classification model for Stock Maret data for 1089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010. Our goal is to predict the trend of the market as "Up" or "Down" given some predictors; returns for Lag 1 to Lag5 & Volume


## Dataset 

```{r}

names(Weekly)
head(Weekly)
attach(Weekly)
cor(Weekly[1:8])

plot(Volume)

```


## Logistic Regression

```{r}

glm.fit=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,family = binomial,data=Weekly)

summary(glm.fit)

contrasts(Direction)

glm.probabilities=predict(glm.fit,type="response")

glm.prediction=rep("Down",dim(Weekly)[1])

glm.prediction[glm.probabilities>0.5]="Up"

table(glm.prediction,Direction)

mean(glm.prediction==Direction)
mean(glm.prediction!=Direction)

```

## Training & Testing Model

```{r}
train=Year<2007

Data.predictors=Weekly[!train,]
Data.response=Direction[!train]


glm.fit=glm(Direction~Lag2,data=Weekly,family=binomial,subset=train)

summary(glm.fit)

glm.probabilities=predict(glm.fit,Data.predictors,type="response")

glm.prediction=rep("Down",length(Data.response))
glm.prediction[glm.probabilities>0.5]="Up"

table(glm.prediction,Data.response)

mean(glm.prediction==Data.response)
mean(glm.prediction!=Data.response)
```


## Linear Discriminant Analysis

```{r}
lda.fit=lda(Direction~Lag2,data=Weekly,subset=train)

lda.fit

plot(lda.fit)

lda.prediction=predict(lda.fit, Data.predictors)
lda.class=lda.prediction$class

table(lda.class,Data.response)

mean(lda.class==Data.response)
mean(lda.class!=Data.response)
```

## Quadratic Discriminant Analysis

```{r}
qda.fit=qda(Direction~Lag2,data=Weekly,subset=train)

qda.fit


qda.prediction=predict(qda.fit, Data.predictors)
qda.class=lda.prediction$class

table(qda.class,Data.response)

mean(qda.class==Data.response)
mean(qda.class!=Data.response)
```

## K-Nearest Neighbors
```{r}

train.X=cbind(Lag1,Lag2)[train,]


test.X=cbind(Lag1,Lag2)[!train,]


train.Direction=Direction[train]


set.seed(1)
knn.pred=knn(train.X,test.X,train.Direction,k=1)
table(knn.pred,Data.response)
mean(knn.pred==Data.response)


knn.pred=knn(train.X,test.X,train.Direction,k=3)
table(knn.pred,Data.response)

mean(knn.pred==Data.response)
```

## Conclusions

We have compared the performance on test data for 2008 to 2010 on different classifiers trained from 1990 to 2007. The classification alrotighms used were Logistic Regression, Linear Discriminant Analysis, Quadratic Distriminant Analysis & K-Nearest Neighbors and they all perform the same, having a correct classification score of 55% which is only slightly better than random guessing thus we can conclude that this dataset is not a great predictor to beat the market
