---
title: "Math 208 Final Project"
author: "Aymen Rumi"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# Task 1: [**Exploratory Single Variable Analysis**](https://github.com/yixuan/prettydoc/)


Using `Womens_Clothing_Reviews.csv` dataset, we will provide some `Exploratory Data Analyses` and describe the distributions of [**age**](https://github.com/yixuan/prettydoc/), [**product rating**](https://github.com/yixuan/prettydoc/), [**recommendations**](https://github.com/yixuan/prettydoc/), and [**article departments**](https://github.com/yixuan/prettydoc/) amongst the respondents




`Functions` we will use include
```yaml
DataVisualization: Produces graphic visual of numerical/categorical data
FrequencyTable: Produces table of variable with counts
SummaryTable: Produces a table showing mean,median,standard deviation & quantiles
```

## [**Function Definition**](https://github.com/yixuan/prettydoc/)

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(knitr)
Womens_Clothing_Review<-read_csv("Womens_Clothing_Reviews.csv")
Womens_Clothing_Review<-Womens_Clothing_Review%>%drop_na()
```

```{r fig.width=6, fig.height=6, fig.align='center'}

DataVisualization<-function(data,subset,numerical=TRUE,name)
{
  
  if(numerical)
  {
    
    ggplot(data,(aes(x=subset)))+geom_bar(fill="lightblue",color="black")+
      scale_fill_viridis_d()+xlab(name)
  }
  else
  {
    ggplot(data,(aes(x=subset,fill=subset)))+
      geom_bar()+scale_fill_viridis_d()+xlab(name)
  }
  
}

```

```{r fig.width=6, fig.height=6, fig.align='center'}

FrequencyTable<-function(data)
{
  data%>%summarise(count=n())%>%mutate(prop=count/sum(count))%>%
    arrange(desc(count))%>%kable()
}

```

```{r fig.width=6, fig.height=6, fig.align='center'}

SummaryTable<-function(data,subset)
{
  data%>%summarise(Ave=mean(subset),
                   Med=median(subset),
                   '25%ile'=quantile(subset,0.25),
                   '75%ile'=quantile(subset,0.75),
                   Std=sd(subset)
                   )%>%kable()
}

```

## [**Variable Analysis:**](https://github.com/yixuan/prettydoc/) Age

The figures below provide the `Density Plot`, `Ordered Count/Proportion Table`, and `Summary Table` for [**age**](https://github.com/yixuan/prettydoc/)
```{r fig.width=6, fig.height=6, fig.align='center'}

DataVisualization(Womens_Clothing_Review,with(Womens_Clothing_Review,Age),name="Age")
FrequencyTable(Womens_Clothing_Review%>%group_by(Age))
SummaryTable(Womens_Clothing_Review,with(Womens_Clothing_Review,Age))

```


`Analysis` 
```yaml
We can see that the average age is approximately 43 with
a spread of 12, the density has a high peak for age
39(1103 people) which is 375 more than the next highest
age group of 35, the 25th and 75th quartile are between
34 and 53
```


## [**Variable Analysis:**](https://github.com/yixuan/prettydoc/) Rating

The figures below provide the `Density Plot` and `Ordered Count/Proportion Table` for [**rating**](https://github.com/yixuan/prettydoc/)

```{r fig.width=6, fig.height=6, fig.align='center'}
DataVisualization(Womens_Clothing_Review,
                  with(Womens_Clothing_Review,Rating),name="Rating")
FrequencyTable(Womens_Clothing_Review%>%group_by(Rating))

```

`Analysis` 
```yaml
Looks like more a little more than half of reviews are positive with 5 stars,
and the rest combine for the other half with 4 taking approximately 20%
3 taking 12% and the rest for 1&2, overall very good ratings
```

## [**Variable Analysis:**](https://github.com/yixuan/prettydoc/) Recommendations

The figures below provide the `Density Plot` and `Ordered Count/Proportion Table` for [**recommendations**](https://github.com/yixuan/prettydoc/)

```{r fig.width=6, fig.height=6, fig.align='center'}

DataVisualization(Womens_Clothing_Review,
                  with(Womens_Clothing_Review,Recommended),name="Recommended")
FrequencyTable(Womens_Clothing_Review%>%group_by(Recommended))
```

`Analysis` 
```yaml
There is a 80% positive recommendation and 20% negative,
overall high recommendation rating

```

## [**Variable Analysis:**](https://github.com/yixuan/prettydoc/) Departments

The figures below provide the `Density Plot` and `Ordered Count/Proportion Table` for [**recommendations**](https://github.com/yixuan/prettydoc/)

```{r fig.width=6, fig.height=6, fig.align='center'}
DataVisualization(Womens_Clothing_Review,with(Womens_Clothing_Review,
                                              Department_Name),FALSE,"Departments")
FrequencyTable(Womens_Clothing_Review%>%
                 group_by(Department_Name))

```

`Analysis` 
```yaml
There is a higher percentage of article clothings that are
tops(44%) with second in dresses(27%) and the
rest of the 3 each getting significantly lower

```



# Task 2: [**Exploring Associations**](https://github.com/yixuan/prettydoc/)

Using `Womens_Clothing_Reviews.csv` , we will explore 2 questions.

` Question 1` We will look at distributons of [**age**](https://github.com/yixuan/prettydoc/) accross [**article departments**](https://github.com/yixuan/prettydoc/)


` Question 2` We will look at five demographic categories: [**25 and under**](https://github.com/yixuan/prettydoc/), [**26 - 35**](https://github.com/yixuan/prettydoc/), [**36-45**](https://github.com/yixuan/prettydoc/), [**46-64**](https://github.com/yixuan/prettydoc/), and [**65 and over**](https://github.com/yixuan/prettydoc/) and compare the distribution of product ratings amongst each of the five age groups



`Functions` we will use include
```yaml
Department_Summary: Produces a frequency table for a specific Department
Age_Category: Takes in int and puts in age category

```

## [**Function Definition**](https://github.com/yixuan/prettydoc/)

```{r fig.width=6, fig.height=6, fig.align='center'}

Department_Summary<-function(department)
{
  Womens_Clothing_Review%>%filter(Department_Name==department)%>%
    summarise(Ave=mean(Age),Med=median(Age),
                    '25%ile'=quantile(Age,0.25),
                    '75%ile'=quantile(Age,0.75),
                    Std=sd(Age))%>%kable()
}

```

```{r fig.width=6, fig.height=6, fig.align='center'}


Age_Category<-function(value)
{
    if(value<=25)
    {
        return("25 and under")
    }
    else if((value>=26)&&(value<=35))
    {
        return("26-35")
    }
    else if((value>=36)&&(value<=45))
    {
        return ("36-45")
    }
    else if((value>=36)&&(value<=64))
    {
        return("46-64")
    }
    else
    {
        return("65 and over")
    }
}
    
 
 
```


## [**Question 1:**](https://github.com/yixuan/prettydoc/)

`Data Visualization`
```{r fig.width=6, fig.height=6, fig.align='center'}
ggplot(Womens_Clothing_Review,aes(Age,fill=Department_Name))+
  geom_bar()+facet_wrap(~Department_Name)

```

`Bottoms: Age Distribution Table`
```{r fig.width=6, fig.height=6, fig.align='center'}
Department_Summary("Bottoms")
```

`Intimate: Age Distribution Table`
```{r fig.width=6, fig.height=6, fig.align='center'}
Department_Summary("Intimate")
```

`Jackets: Age Distribution Table`
```{r fig.width=6, fig.height=6, fig.align='center'}
Department_Summary("Jackets")
```

`Tops: Age Distribution Table`
```{r fig.width=6, fig.height=6, fig.align='center'}
Department_Summary("Tops")
```

`Trends: Age Distribution Table`
```{r fig.width=6, fig.height=6, fig.align='center'}
Department_Summary("Trend")

```



`Analysis` 
```yaml
The distribution as in the spread of age seems to look relatively the same across
all departments but the densities differ greatly,the density of
people rating products are in order of Tops, Dresses, Jackets,
Intimate,Jackets, Trends

```

## [**Question 2:**](https://github.com/yixuan/prettydoc/)

`Data Visualization`
```{r fig.width=6, fig.height=6, fig.align='center'}

Womens_Clothing_Review<-Womens_Clothing_Review%>%
  mutate(Category=map_chr(Age,Age_Category))


ggplot(Womens_Clothing_Review,aes(x=Rating,fill=Category))+
  geom_bar()+facet_wrap(~Category)

```


`Analysis` 

```yaml
Again the distribution as in the spread of rating seems to look 
relatively the same across all age groups but the densities 
differ greatly due to the fact that there is more counts of 
ratings from certain groups, groups such as 46-64 and 36-45 
age group have a vast amount of ratings and seem to be most 
enthusiastic about their company’s products.
```


## Task 3: [**Ten Most Popular Products**](https://github.com/yixuan/prettydoc/)

Using `Womens_Clothing_Reviews.csv` dataset, we will compile a list of their [**ten most popular products**](https://github.com/yixuan/prettydoc/) based on [**Recommendations**](https://github.com/yixuan/prettydoc/) (with each product indicated by [**ID number**](https://github.com/yixuan/prettydoc/))

We will use `Wilson’s Lower Confidence Limit` computed via: 

[**n**](https://github.com/yixuan/prettydoc/) = the `number` of respondents who rated that product (positively or negatively)

[**p**](https://github.com/yixuan/prettydoc/)=the `proportion` of respondents who positively recommended a certain product

[*$a = \frac{1.96^2}{n}$*](https://github.com/yixuan/prettydoc/)

[*$b = \frac{p(1-p)}{n}$*](https://github.com/yixuan/prettydoc/)

[*$c = \frac{a}{2n}$*](https://github.com/yixuan/prettydoc/)

[*$WLCL = \frac{p+a-1.96\sqrt{b+c}}{1+2a}$*](https://github.com/yixuan/prettydoc/)


`Functions` we will use include
```yaml
Clothing: returns the proportion of positively recommended ratings for given ID
Compute_WLCL: Given a dataset will compute its WLCL and output it

```

## [**Function Definition**](https://github.com/yixuan/prettydoc/)

```{r fig.width=6, fig.height=6, fig.align='center'}
Clothing<-function(ID)
{

    max((Womens_Clothing_Review%>%filter(Clothing_ID%in%ID)%>%
           group_by(Recommended)%>%summarise(count=n()))%>%
          mutate(prop=count/sum(count))%>%select(prop))
    
}
```

```{r fig.width=6, fig.height=6, fig.align='center'}

Compute_WLCL<-function(Womens_Clothing_Review)
{
  
  Womens_Clothing_Review<-Womens_Clothing_Review%>%
    group_by(Clothing_ID)%>%mutate(prop=map_dbl(Clothing_ID,Clothing))
  Womens_Clothing_Review<-Womens_Clothing_Review%>%
    mutate(a=(1.96**2)/(2*n))
  Womens_Clothing_Review<-Womens_Clothing_Review%>%
    mutate(b=(prop*(1-prop))/(n))
  Womens_Clothing_Review<-Womens_Clothing_Review%>%
    mutate(c=(a)/(2*n))
  Womens_Clothing_Review<-Womens_Clothing_Review%>%
    mutate(WLCL=(prop+a-1.96*sqrt(b+c))/(1+2*a))

  head(distinct(Womens_Clothing_Review%>%
                  arrange(desc(WLCL))%>%select(Clothing_ID,WLCL,n,Department_Name)),10)%>%kable()
}
```


## [**Part A):**](https://github.com/yixuan/prettydoc/)
`the 10 product ID’s with the highest average ratings`

```{r fig.width=6, fig.height=6, fig.align='center'}
Womens_Clothing_Review<-Womens_Clothing_Review%>%
  group_by(Clothing_ID)%>%mutate(n=n())

head(Womens_Clothing_Review%>%group_by(Clothing_ID)%>%
       mutate(Mean_Rating= mean(Rating))%>%select(Clothing_ID,n,Department_Name,Mean_Rating)%>%
       arrange(desc(Mean_Rating)),10)%>%kable()
```


## [**Part B):**](https://github.com/yixuan/prettydoc/)
`the 10 product ID’s with the highest proportion of positive recommendations`

```{r fig.width=6, fig.height=6, fig.align='center'}
head(Womens_Clothing_Review%>%group_by(Clothing_ID)%>%
       mutate(prop=map_dbl(Clothing_ID,Clothing))%>%
       select(Clothing_ID,prop,n,Department_Name)%>%
       arrange(desc(prop))%>%distinct(),10)%>%kable()
```

## [**Part C):**](https://github.com/yixuan/prettydoc/)
`the 10 product ID’s with the highest Wilson lower confidence limits`

```{r fig.width=6, fig.height=6, fig.align='center'}
Compute_WLCL(Womens_Clothing_Review)

```


`Analysis` 

```yaml
I think the list that best represents the products which are most popular are,
List C for WLCL because the lists shown in A and B can be deceiving as the 
number of counts can be low and this matters for showing an indication of popularity and
i think that WLCL captures that well, as opposed to showing just rating and positive
recommendation proportions it measures the popularity 
```










